---
title: "Post-MIMS concentration and flux calculations for flow-through core experiments"
author: "O'Conall Eller, Colin G. Finlay"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
---

# Thinking about for later: where do we want to reference Hamme, other flux equation authors, and AR Smyth calculation sheet as source material?

##MIMS Flow
What you need:
- MIMS Data .xlsx
- Flow Rates .xlsx

What you need to change ( * ):
- Hamme Solubility
  - Calculator
    - Insert
*      - Change Temperature and Salinity Values for stds, first set, and second,
set of calculations. 

- MIMS Data
*  - Insert MIMS Data .xlsx
*  - Change Sheet name and data range (if needed)

- MIMS Calculations
*  - Row data ranges (ex. 56:91)

- MIMS First Set of Samples (Fresh, etc.)
*  - Row data ranges (ex. 1:45)
*  - Optional sections of code at the bottom

- MIMS Second Set of Samples (Salt, etc.)
*  - Row data ranges (ex. 72:107)
*  - Optional sections of code at the bottom

- Flux Calculations
  - Flow Rates
*    - Insert Flow Rates .xlsx
*    - Change Sheet name and data range (if needed)
*      - Change data ranges (ex. 1:24, 26:49)
      
- First Set of Flux Calculations (Fresh, etc.)
*  - Change data ranges (ex. 34:36, 1:36 )
*  - Change Area value

- Second Set of Flux Calculations (Salt, etc.)
*  - Change data ranges (ex. 34:36, 37:72)
*  - Change Area value

- Verification
  - Cleaning up Excel
*    - Adjust BP rows to be removed
  - Make sure R and Excel Data are aligned
  - Input
*    - Assign data_frame, excel_file, sheet_name, and data_range.


Notes:
- May need to change "sample" to "sample ID" (or vise versa) depending on your
MIMS Data
- May need to change "Time" to "time run" (or vise versa) depending on your 
MIMS Data

##Housekeeping

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
#use to set working directory 
knitr::opts_knit$set(root.dir="~/GitHub/FlowThrough_CoastalAg/MIMS_calculations/")
```

```{r load packages}
library(readxl)
library(dplyr)
library(tidyr)
```

##Hamme Solubility -----

##Table of Coefficients
```{r}
coef_table <- data.frame(
  Coefficient = c("A0", "A1", "A2", "A3", "A4", "A5",
                  "B0", "B1", "B2", "B3", "C0", "XcSTD"),
  
  N2 = c(6.42931, 2.92704, 4.32531, 4.69149,
         NA, NA,
         -0.00744129, -0.00802566, -0.0146775,
         NA, NA,
         22.404),
 Ar = c(2.7915, 3.17609, 4.13116, 4.90379,
         NA, NA,
         -0.00696233, -0.0076667, -0.0116888,
         NA, NA,
         22.393),
  
  Ne = c(2.18156, 1.29108, 2.12504, 0,
         NA, NA,
         -0.00594737, -0.00513896, 0,
         NA, NA,
         22.425),
  
  O2 = c(5.8087100, 3.2029100, 4.1788700, 5.1000600,
         -0.0986643, 3.8036900,
         -0.00701577, -0.00770028, -0.0113864,
         -0.00951519, -0.000000275915,
         NA))

```


## Calculator
```{r}
calc_sample <- function(temp, sal) {
  ##Ts, Ts2, Ts3 (Ts4, Ts5 (not shown)) Calculation 
  Ts <- log((298.15 - temp) / (273.15 + temp))
  Ts2 <- Ts^2
  Ts3 <- Ts^3
  Ts4 <- Ts3 * Ts
  Ts5 <- Ts3 * Ts2
 
  ##[N2] Calculation
  A0 <- coef_table$N2[coef_table$Coefficient == "A0"]
  A1 <- coef_table$N2[coef_table$Coefficient == "A1"]
  A2 <- coef_table$N2[coef_table$Coefficient == "A2"]
  A3 <- coef_table$N2[coef_table$Coefficient == "A3"]
  
  B0 <- coef_table$N2[coef_table$Coefficient == "B0"]
  B1 <- coef_table$N2[coef_table$Coefficient == "B1"]
  B2 <- coef_table$N2[coef_table$Coefficient == "B2"]
  
  ln_N2 <-
    A0 + (A1 * Ts) + (A2 * Ts2) + (A3 * Ts3) + 
    (sal * (B0 + (B1 * Ts) + (B2 * Ts2)))
  
  N2 <- exp(ln_N2)
  
  ##28N2 Calculation
  N2_28 <- N2 * 0.99634 * 0.99634
  
  ##29N2 Calculation
  N2_29 <- N2 * 0.0072932
  
  ##[Ar] Calculation
  A0 <- coef_table$Ar[coef_table$Coefficient == "A0"]
  A1 <- coef_table$Ar[coef_table$Coefficient == "A1"]
  A2 <- coef_table$Ar[coef_table$Coefficient == "A2"]
  A3 <- coef_table$Ar[coef_table$Coefficient == "A3"]

  B0 <- coef_table$Ar[coef_table$Coefficient == "B0"]
  B1 <- coef_table$Ar[coef_table$Coefficient == "B1"]
  B2 <- coef_table$Ar[coef_table$Coefficient == "B2"]
  
  ln_Ar <- 
    (A0 + A1*Ts + A2*Ts2 + A3*Ts3 + sal*(B0 + B1*Ts + B2*Ts2))
  
  Ar <- exp(ln_Ar)
    
  ##40Ar Calculation
  Ar_40 <- Ar * 0.996
  
  ##28N2/40Ar Calculation
  ratio_N2_Ar <- N2_28 / Ar_40
  
  ##29N2/40Ar
  ratio_29_Ar <- N2_29 / Ar_40
  
  ## [Ne] Calculation
  A0 <- coef_table$Ne[coef_table$Coefficient == "A0"]
  A1 <- coef_table$Ne[coef_table$Coefficient == "A1"]
  A2 <- coef_table$Ne[coef_table$Coefficient == "A2"]
  A3 <- coef_table$Ne[coef_table$Coefficient == "A3"]
  
  B0 <- coef_table$Ne[coef_table$Coefficient == "B0"]
  B1 <- coef_table$Ne[coef_table$Coefficient == "B1"]
  B2 <- coef_table$Ne[coef_table$Coefficient == "B2"]
  
  ln_Ne <-
    (A0 + (A1 * Ts) + (A2 * Ts2) + (A3 * Ts3) + (sal * 
          (B0 + (B1 * Ts) + (B2 * Ts2))))
  
  Ne <- exp(ln_Ne)
  
  ##[O2] Calculation
  A0 <- coef_table$O2[coef_table$Coefficient == "A0"]
  A1 <- coef_table$O2[coef_table$Coefficient == "A1"]
  A2 <- coef_table$O2[coef_table$Coefficient == "A2"]
  A3 <- coef_table$O2[coef_table$Coefficient == "A3"]
  A4 <- coef_table$O2[coef_table$Coefficient == "A4"]
  A5 <- coef_table$O2[coef_table$Coefficient == "A5"]
  
  B0 <- coef_table$O2[coef_table$Coefficient == "B0"]
  B1 <- coef_table$O2[coef_table$Coefficient == "B1"]
  B2 <- coef_table$O2[coef_table$Coefficient == "B2"]
  B3 <- coef_table$O2[coef_table$Coefficient == "B3"]
  C0 <- coef_table$O2[coef_table$Coefficient == "C0"]
  
  ln_O2 <-
    A0 + A1 * Ts + A2 * Ts2 + A3 * Ts3 + A4 * Ts4 + A5 * Ts5 + 
    sal * (B0 + B1 * Ts + B2 * Ts2 + B3 * Ts3 + C0 * Ts2)
  
  O2 <- exp(ln_O2)
  
  ##32O2 Calculation
  O2_32 <- O2 * 0.99759 * 0.99759
  
  ##32O2/40Ar Calculation
  ratio_O232_Ar <- O2_32 / Ar_40
  
  data.frame(
    TEMP = temp,
    SALINITY = sal,
    Ts = Ts,
    Ts2 = Ts2,
    Ts3 = Ts3,
    N2 = N2,
    N2_28 = N2_28,
    N2_29 = N2_29,
    Ar = Ar,
    Ar_40 = Ar_40,
    N2_28_to_Ar_40 = ratio_N2_Ar,
    N2_29_to_Ar_40 = ratio_29_Ar,
    Ne = Ne,
    O2 = O2,
  O2_32 = O2_32,
  O2_32_to_Ar_40 = ratio_O232_Ar)
}

##For umoles and nmoles/liter -----
calc_umoles <- function(temp, sal) {
  
  ##Calling
    sample <- calc_sample(temp, sal)
    N2 <- sample$N2
    N2_28 <- sample$N2_28
    N2_29 <- sample$N2_29
    Ar <- sample$Ar
    Ar_40 <- sample$Ar_40
    Ne <- sample$Ne
    O2 <- sample$O2
    O2_32 <- sample$O2_32
    ratio_O232_Ar <- sample$ratio_O232_Ar
    ratio_N2_Ar <- sample$ratio_N2_Ar
    ratio_29_Ar <- sample$ratio_29_Ar
    
    
  
  ##For density calculations
  DCA <- 0.824493 - 0.0040899 * temp + 0.000076438 * (temp^2) - 0.00000082467 * (temp^3) + 0.0000000053875 * (temp^4)
  DCB <- -0.00572466 + 0.00010227 * temp - 0.0000016546 * (temp^2)
  DCC <- 0.00048314
  
  
  ##po and p calculations
  po <- 999.842594 + (0.06793952) * (temp)  - ( 0.00909529) * (temp^2) +
  (0.0001001685) * (temp^3) - (0.000001120083) * (temp^4) + (0.000000006536336) * (temp^5)
  
  p <- po + (DCA*sal)+((DCB*sal)^3/2) + DCC*sal
  
  p_kg_l <- p / 1000
  
  ## N2 Calculations
  
  uM_N2 <- N2 * p_kg_l
  
  uM_N2_28 <- N2_28 * p_kg_l
  
  uM_N2_29 <- N2_29 * p_kg_l
  
  ##Ar Calculations
  
  uM_Ar <- Ar * p_kg_l
  
  uM_40_Ar <- Ar_40 * p_kg_l
  
  ##ratios
  uM_ratio_N2_Ar <- uM_N2_28 / uM_40_Ar
  
  uM_ratio_29_Ar <- uM_N2_29 / uM_40_Ar
  
  ##Ne Calculations
  
  uM_Ne <- Ne * p_kg_l
  
  ##O2 Calculations
  
  uM_O2 <- O2 * p_kg_l
  
  uM_O2_32 <- O2_32 * p_kg_l
  
  uM_ratio_O232_Ar <- uM_O2_32 / uM_40_Ar
  
  
  data.frame(
    TEMP = temp,
    SALINITY = sal,
    po = po,
    p = p,
    p_kg_l = p_kg_l,
    uM_N2 = uM_N2,
    uM_N2_28 = uM_N2_28,
    uM_N2_29 = uM_N2_29,
    uM_Ar = uM_Ar,
    uM_40_Ar = uM_40_Ar,
    uM_N2_28_to_Ar_40 = uM_ratio_N2_Ar,
    uM_N2_29_to_Ar_40 = uM_ratio_29_Ar,
    uM_Ne = uM_Ne,
    uM_O2 = uM_O2,
    uM_O2_32 = uM_O2_32,
    uM_O2_32_to_Ar_40 = uM_ratio_O232_Ar
    
  )

}

##Insert (sample temp, sample sal) -----
gas_ratios_stds <- calc_sample(14, 0)# <- Not used in calculations
uM_nM_stds <- calc_umoles(14,0) # <-- For stds calculations
uM_nM_fresh <- calc_umoles(14,20) # <-- For first set of data 
uM_nM_salt <- calc_umoles(14,21.16) # <- For second set of data (may not be needed)

#If any renamed, don't forget to replace all df names in future calculations
```

##MIMS Data -----
```{r}
# Colin comment: add any input data files to "../data/"
  # MRW: "../data/Hyde/MRW/"
  # YHR (June flow): "../data/Pitt/YHR/"
# Colin comment: read_excel is new to me. If it works well we can stick with it. If it's clunky, we can easily switch to .csv input data and read.csv

# Previous
#MIMS_Data <- read_excel("~/Downloads/ECU/Research/Peralta_Lab/Coding_Project/data/20260521_Finlay_MIMS_72Samples_ARScheck (copy).xlsx",
                        
                  #sheet = "Data",
                   #range = "A8:L115") # <- Adjust accordingly

# Updated, date file in repo, relative path
MIMS_Data <- read_excel("../data/Hyde/MRW/20260521_Finlay_MIMS_72Samples_ARScheck.xlsx",
                        
                  sheet = "Data",
                   range = "A8:L115") # <- Adjust accordingly

#Depending on the name, you may need to adjust "sample" to "sample ID"

```

##MIMS Calculations -----
```{r}
##MIMS Drift -----
#Row numbers should be your stds rows
MIMS_Data$`N2:Ar` <- as.numeric(MIMS_Data$`N2:Ar`)
MIMS_Data$`time run` <- as.numeric(MIMS_Data$`time run`)
MIMS_Data$`O2:Ar` <- as.numeric(MIMS_Data$`O2:Ar`)
MIMS_Data$`40` <- as.numeric(MIMS_Data$`40`)

MIMS_Drift_df <- data.frame(
  metric = c("Slope", "Intercept"),

  N2_Ar = c(
    coef(lm(`N2:Ar` ~ `time run`, data = MIMS_Data[1:33, ]))[2],
    coef(lm(`N2:Ar` ~ `time run`, data = MIMS_Data[1:33, ]))[1]
  ),

  O2_Ar = c(
    coef(lm(`O2:Ar` ~ `time run`, data = MIMS_Data[1:33, ]))[2],
    coef(lm(`O2:Ar` ~ `time run`, data = MIMS_Data[1:33, ]))[1]
  ),
  
  Ar = c(
    coef(lm(`40` ~ `time run`, data = MIMS_Data[1:33, ]))[2],
    coef(lm(`40` ~ `time run`, data = MIMS_Data[1:33, ]))[1]
  )
)
## STDS Calculations -----

#N2:Ar stnd
slope_N2 <- MIMS_Drift_df$N2_Ar[MIMS_Drift_df$metric == "Slope"]
intercept_N2 <- MIMS_Drift_df$N2_Ar[MIMS_Drift_df$metric == "Intercept"]

#O2:Ar stnd
slope_O2 <- MIMS_Drift_df$O2_Ar[MIMS_Drift_df$metric == "Slope"]
intercept_O2 <- MIMS_Drift_df$O2_Ar[MIMS_Drift_df$metric == "Intercept"]

#[uM N2]:[uM Ar] (Stds Water Bath Equilibrium)
uM_N2_uM_Ar_value <- uM_nM_stds$`uM_N2_28_to_Ar_40`

#Ar stnd
slope_Ar <- MIMS_Drift_df$Ar[MIMS_Drift_df$metric == "Slope"]
intercept_Ar <- MIMS_Drift_df$Ar[MIMS_Drift_df$metric == "Intercept"]

#[uM O2]:[uM Ar] (water bath equilibrium)
uM_O2_uM_Ar_value <- uM_nM_stds$`uM_O2_32_to_Ar_40`

#[uM N2]:[uM Ar] (sample)
x <- MIMS_Data$`time run`[1:33]

N2_Ar_stnd = slope_N2 * x + intercept_N2
  
uM_N2_uM_Ar_sample_value <-
  (MIMS_Data$`N2:Ar`[1:33] * uM_N2_uM_Ar_value) /
  N2_Ar_stnd

#[uM O2]:[uM Ar] (sample)
O2_Ar_stnd = slope_O2 * x + intercept_O2

uM_O2_uM_Ar_sample_value <-
  (MIMS_Data$`O2:Ar`[1:33] * uM_O2_uM_Ar_value) /
  O2_Ar_stnd

#uM Ar (exp equilibrium)
uM_Ar_exp_value <- uM_nM_stds$`uM_40_Ar`

#uM N
uM_N_value <- (uM_N2_uM_Ar_sample_value * uM_Ar_exp_value) * 2

#uM O2
uM_O2_value <- (uM_O2_uM_Ar_sample_value * uM_Ar_exp_value)

#mg/L O2
mg_L_O2_value <- (uM_O2_value * 32) / 1000

#uM Ar
Ar_stnd <- slope_Ar * x + intercept_Ar

uM_Ar_value <- (MIMS_Data$`40`[1:33]) * (uM_Ar_exp_value / Ar_stnd)

#2STDEV_uMN
STDEV_2_uMN_value = 2 * (sd(uM_N_value))

#2STDEV_uMO2
STDEV_2_uMO2_value = 2 * (sd(uM_O2_value))

#2STDEV_uMAr
STDEV_2_uMAr_value = 2 * (sd(uM_Ar_value))


MIMS_stds_Calc <- data.frame(
  sample = "stds",
  N2_Ar_stnd = slope_N2 * x + intercept_N2,
  O2_Ar_stnd = slope_O2 * x + intercept_O2,
  uM_N2_uM_Ar = uM_N2_uM_Ar_value,
  Ar_stnd = slope_Ar * x + intercept_Ar,
  uM_O2_uM_Ar = uM_O2_uM_Ar_value,
  uM_N2_uM_Ar_sample = uM_N2_uM_Ar_sample_value,
  uM_O2_uM_Ar_sample = uM_O2_uM_Ar_sample_value,
  uM_Ar_exp = uM_Ar_exp_value,
  uM_N = uM_N_value,
  uM_O2 = uM_O2_value,
  mg_L_O2 = mg_L_O2_value,
  uM_Ar = uM_Ar_value,
  STDEV_2_uMN = STDEV_2_uMN_value,
  STDEV_2_uMO2 = STDEV_2_uMO2_value,
  STDEV_2_uMAr = STDEV_2_uMAr_value

)
```

##MIMS First Set of Samples (Fresh, etc.)
```{r}
## First Set of Samples Calculations

#N2:Ar stnd
slope_N2 <- MIMS_Drift_df$N2_Ar[MIMS_Drift_df$metric == "Slope"]
intercept_N2 <- MIMS_Drift_df$N2_Ar[MIMS_Drift_df$metric == "Intercept"]

#O2:Ar stnd
slope_O2 <- MIMS_Drift_df$O2_Ar[MIMS_Drift_df$metric == "Slope"]
intercept_O2 <- MIMS_Drift_df$O2_Ar[MIMS_Drift_df$metric == "Intercept"]

#[uM N2]:[uM Ar] (Stds Water Bath Equilibrium)
uM_N2_uM_Ar_value <- uM_nM_stds$`uM_N2_28_to_Ar_40`

#Ar stnd
slope_Ar <- MIMS_Drift_df$Ar[MIMS_Drift_df$metric == "Slope"]
intercept_Ar <- MIMS_Drift_df$Ar[MIMS_Drift_df$metric == "Intercept"]

#[uM O2]:[uM Ar] (water bath equilibrium)
uM_O2_uM_Ar_value <- uM_nM_stds$`uM_O2_32_to_Ar_40`

#[uM N2]:[uM Ar] (sample)
x <- MIMS_Data$`time run`[36:71]

N2_Ar_stnd = slope_N2 * x + intercept_N2
  
uM_N2_uM_Ar_sample_value <-
  (MIMS_Data$`N2:Ar`[36:71] * uM_N2_uM_Ar_value) /
  N2_Ar_stnd

#[uM O2]:[uM Ar] (sample)
O2_Ar_stnd = slope_O2 * x + intercept_O2

uM_O2_uM_Ar_sample_value <-
  (MIMS_Data$`O2:Ar`[36:71] * uM_O2_uM_Ar_value) /
  O2_Ar_stnd

#uM Ar (exp equilibrium)
uM_Ar_exp_value <- uM_nM_fresh$`uM_40_Ar`

#uM N
uM_N_value <- (uM_N2_uM_Ar_sample_value * uM_Ar_exp_value) * 2

#uM O2
uM_O2_value <- (uM_O2_uM_Ar_sample_value * uM_Ar_exp_value)

#mg/L O2
mg_L_O2_value <- (uM_O2_value * 32) / 1000

#uM Ar
Ar_stnd <- slope_Ar * x + intercept_Ar

uM_Ar_value <- (MIMS_Data$`40`[36:71]) * (uM_Ar_exp_value / Ar_stnd)


#Dataframe (adjust name as needed)
MIMS_fresh_Calc <- data.frame(
  `sample` = MIMS_Data$`sample`[36:71],
  N2_Ar_stnd = slope_N2 * x + intercept_N2,
  O2_Ar_stnd = slope_O2 * x + intercept_O2,
  Ar_stnd = slope_Ar * x + intercept_Ar,
  uM_N2_uM_Ar_stds = uM_N2_uM_Ar_value,
  uM_O2_uM_Ar_stds = uM_O2_uM_Ar_value,
  uM_N2_uM_Ar_sample = uM_N2_uM_Ar_sample_value,
  uM_O2_uM_Ar_sample = uM_O2_uM_Ar_sample_value,
  uM_Ar_exp = uM_Ar_exp_value,
  uM_N = uM_N_value,
  uM_O2 = uM_O2_value,
  mg_L_O2 = mg_L_O2_value,
  uM_Ar = uM_Ar_value
)

MIMS_fresh_BP <- MIMS_fresh_Calc[c(34, 35, 36), ] # <-- Creating a seperate df for BP rows to make future calculations easier (change to match dataset)

# MIMS_fresh_BP <- MIMS_fresh_BP[-c(1, 4), ] # <-- ONLY IF you need to remove some BP rows

MIMS_fresh_Calc <- MIMS_fresh_Calc[-c(34, 35, 36), ] # <-- To remove BP rows
```

## MIMS Second Set of Samples (Salt, etc.)
```{r}
## Second Set of Samples Calculations
# You may not need to run this chunk of code if you only are using one set of data that does not differ 

#N2:Ar stnd
slope_N2 <- MIMS_Drift_df$N2_Ar[MIMS_Drift_df$metric == "Slope"]
intercept_N2 <- MIMS_Drift_df$N2_Ar[MIMS_Drift_df$metric == "Intercept"]

#O2:Ar stnd
slope_O2 <- MIMS_Drift_df$O2_Ar[MIMS_Drift_df$metric == "Slope"]
intercept_O2 <- MIMS_Drift_df$O2_Ar[MIMS_Drift_df$metric == "Intercept"]

#[uM N2]:[uM Ar] (Stds Water Bath Equilibrium)
uM_N2_uM_Ar_value <- uM_nM_stds$`uM_N2_28_to_Ar_40`

#Ar stnd
slope_Ar <- MIMS_Drift_df$Ar[MIMS_Drift_df$metric == "Slope"]
intercept_Ar <- MIMS_Drift_df$Ar[MIMS_Drift_df$metric == "Intercept"]

#[uM O2]:[uM Ar] (water bath equilibrium)
uM_O2_uM_Ar_value <- uM_nM_stds$`uM_O2_32_to_Ar_40`

#[uM N2]:[uM Ar] (sample)
x <- MIMS_Data$`time run`[72:107]

N2_Ar_stnd = slope_N2 * x + intercept_N2
  
uM_N2_uM_Ar_sample_value <-
  (MIMS_Data$`N2:Ar`[72:107] * uM_N2_uM_Ar_value) /
  N2_Ar_stnd

#[uM O2]:[uM Ar] (sample)
O2_Ar_stnd = slope_O2 * x + intercept_O2

uM_O2_uM_Ar_sample_value <-
  (MIMS_Data$`O2:Ar`[72:107] * uM_O2_uM_Ar_value) /
  O2_Ar_stnd

#uM Ar (exp equilibrium)
uM_Ar_exp_value <- uM_nM_salt$`uM_40_Ar`

#uM N
uM_N_value <- (uM_N2_uM_Ar_sample_value * uM_Ar_exp_value) * 2

#uM O2
uM_O2_value <- (uM_O2_uM_Ar_sample_value * uM_Ar_exp_value)

#mg/L O2
mg_L_O2_value <- (uM_O2_value * 32) / 1000

#uM Ar
Ar_stnd <- slope_Ar * x + intercept_Ar

uM_Ar_value <- (MIMS_Data$`40`[72:107]) * (uM_Ar_exp_value / Ar_stnd)


#Dataframe (adjust name as needed)
MIMS_salt_Calc <- data.frame(
  `sample` = MIMS_Data$`sample`[72:107],
  N2_Ar_stnd = slope_N2 * x + intercept_N2,
  O2_Ar_stnd = slope_O2 * x + intercept_O2,
  Ar_stnd = slope_Ar * x + intercept_Ar,
  uM_N2_uM_Ar_stds = uM_N2_uM_Ar_value,
  uM_O2_uM_Ar_stds = uM_O2_uM_Ar_value,
  uM_N2_uM_Ar_sample = uM_N2_uM_Ar_sample_value,
  uM_O2_uM_Ar_sample = uM_O2_uM_Ar_sample_value,
  uM_Ar_exp = uM_Ar_exp_value,
  uM_N = uM_N_value,
  uM_O2 = uM_O2_value,
  mg_L_O2 = mg_L_O2_value,
  uM_Ar = uM_Ar_value
)

MIMS_salt_BP <- MIMS_salt_Calc[c(1, 2, 3), ] # <-- Creating a separate df for BP rows to make future calculations easier (change to match dataset)

### MIMS_salt_BP <- MIMS_salt_BP[-c(1, 4), ] # <-- ONLY IF you need to remove some BP rows

MIMS_salt_Calc <- MIMS_salt_Calc[-c(1, 2, 3), ] # <-- To remove BP rows

```

##Flux Calculations
You only need to run one Flow Rate code chunk. Flow Rate (Averaged) is if you need to average the flow rates for the core samples. Flow Rate (Individual) is if you already have all of the flow rates for each core and can input them.

```{r} 
#Flow Rate (Averaged) -----
# Old file path
#Flow_Rates <- read_excel("~/Downloads/ECU/Research/Peralta_Lab/Coding_Project/data/Copy_of_MRW_FlowRates_LineIDs.xlsx",
          # sheet = "June17",
          # range ="A1:U50")

# Updated file path
Flow_Rates <- read_excel("../data/Hyde/MRW/MRW_FlowRates_LineIDs.xlsx",
           #sheet = "AprMay2026",
           range ="AprMay2026!A1:U50")
# NOT LOADING CORRECT SHEET. MAYBE MY FORMATTING ON SHEET AFTER VERSION WAS SHARED. CAN O'CONALL SUPPLY VERSION OF SHEET HE IS USING?

Flow_Rates$Flow_rate_L_hr <- as.numeric(Flow_Rates$Flow_rate_L_hr)

#Flow Rate Averages
flow_rate_value <-(
Flow_Rates$Flow_rate_L_hr[1:24] +
 Flow_Rates$Flow_rate_L_hr[26:49]) / 2

flow_rate_expanded <- rep(flow_rate_value, each =3) # <-- Replicate so we can align the flow rate averages with the samples (one average rate per three core samples)


Flow_Rates_df <- data.frame (
Core_ID = MIMS_Data$sample[36:107], #<- This if you need to add samples to the dataframe if the imported dataset doesnt already contain them.
Flow_rate_L_hr = flow_rate_expanded
)

Flow_Rates_df <- Flow_Rates_df[-c(34, 35, 36, 37, 38, 39), ] # <-- This is to exclude the BP rows to make future calculations easier

```

```{r}
#Flow Rate (Individual) -----
Flow_Rates <- read_excel("~/Downloads/ECU/Research/Peralta_Lab/Coding_Project/data/FlowRates_LineIDs.xlsx", # O'CONALL, PLEASE ADD YOUR VERSION OF THIS FILE TO THE REPO AS WELL."
           sheet = "June11",
           range ="A1:T39")

Flow_Rates$Flow_rate_L_hr <- as.numeric(Flow_Rates$Flow_rate_L_hr)

Flow_Rates_df <- data.frame (
### Core_ID = MIMS_Data$sample[36:107], #<- This if you need to add samples to the dataframe if the imported data set doesn't already contain them.
Flow_rate_L_hr = Flow_Rates$Flow_rate_L_hr)

Flow_Rates_df <- Flow_Rates_df[-c(4, 12, 17, 25, 30, 38), ] # <-- This is to exclude the BP rows to make future calculations easier


```


```{r}
## First Set of Flux Calculations (Fresh, etc) -----

# Note: Row numbers should represent either all of the data or a portion of it (such as the Fresh water portion)

##CV Calculations
CV_N2_Ar_value <- (sd(MIMS_fresh_BP$`uM_N2_uM_Ar_sample`)) / mean(MIMS_fresh_BP$`uM_N2_uM_Ar_sample`)

CV_O2_Ar_value <- (sd(MIMS_fresh_BP$`uM_O2_uM_Ar_sample`)) / mean(MIMS_fresh_BP$`uM_O2_uM_Ar_sample`)

CV_Ar_value <- (sd(MIMS_fresh_BP$`uM_Ar`)) / mean(MIMS_fresh_BP$`uM_Ar`)

Area_m2_value = 0.00311724531052447 # <-- Change Area Value if needed

##Fluxes
N2_flux_value <- ((MIMS_fresh_Calc$`uM_N`) - (mean(MIMS_fresh_BP$`uM_N`))) * (Flow_Rates_df$Flow_rate_L_hr[1:33] / Area_m2_value)

O2_flux_value <- ((MIMS_fresh_Calc$`uM_O2`) - (mean(MIMS_fresh_BP$`uM_O2`))) * (Flow_Rates_df$Flow_rate_L_hr[1:33] / Area_m2_value)

Ar_flux_value <- ((MIMS_fresh_Calc$`uM_Ar`) - (mean(MIMS_fresh_BP$`uM_Ar`))) * (Flow_Rates_df$Flow_rate_L_hr[1:33] / Area_m2_value)

Flux_Fresh <- data.frame(
  sample = Flow_Rates_df$`Core_ID`[1:33],
  CV_N2_Ar = paste0(round(CV_N2_Ar_value * 100, 2), "%"),
  CV_O2_Ar = paste0(round(CV_O2_Ar_value * 100, 2), "%"),
  CV_Ar = paste0(round(CV_Ar_value * 100, 2), "%"),
  N2_flux_umol_N_m2_hr_1 = N2_flux_value,
  O2_flux_umol_N_m2_hr_1 = O2_flux_value,
  Ar_flux_umol_N_m2_hr_1 = Ar_flux_value,
  Flow_L_hr_1 = Flow_Rates_df$Flow_rate_L_hr [1:33],
  Area_m2 = Area_m2_value
)

```

```{r}
## Second Set of Flux Calculations (Salt, etc) -----
# Note: Row numbers should represent either all of the data or a portion of it (such as the Salt water portion)

##CV Calculations
CV_N2_Ar_value <- (sd(MIMS_salt_BP$`uM_N2_uM_Ar_sample`)) / mean(MIMS_salt_BP$`uM_N2_uM_Ar_sample`)

CV_O2_Ar_value <- (sd(MIMS_salt_BP$`uM_O2_uM_Ar_sample`)) / mean(MIMS_salt_BP$`uM_O2_uM_Ar_sample`)

CV_Ar_value <- (sd(MIMS_salt_BP$`uM_Ar`)) / mean(MIMS_salt_BP$`uM_Ar`)

Area_m2_value = 0.00311724531052447 # <-- Change Area Value

##Flux Calculations
N2_flux_value <- ((MIMS_salt_Calc$`uM_N`) - (mean(MIMS_salt_BP$`uM_N`))) * (Flow_Rates_df$Flow_rate_L_hr[34:66] / Area_m2_value)

O2_flux_value <- ((MIMS_salt_Calc$`uM_O2`) - (mean(MIMS_salt_BP$`uM_O2`))) * (Flow_Rates_df$Flow_rate_L_hr[34:66] / Area_m2_value)

Ar_flux_value <- ((MIMS_salt_Calc$`uM_Ar`) - (mean(MIMS_salt_BP$`uM_Ar`))) * (Flow_Rates_df$Flow_rate_L_hr[34:66] / Area_m2_value)

Flux_Salt <- data.frame(
 sample = Flow_Rates_df$`Core_ID`[34:66],
  CV_N2_Ar = paste0(round(CV_N2_Ar_value * 100, 2), "%"),
  CV_O2_Ar = paste0(round(CV_O2_Ar_value * 100, 2), "%"),
  CV_Ar = paste0(round(CV_Ar_value * 100, 2), "%"),
  N2_flux_umol_N_m2_hr_1 = N2_flux_value,
  O2_flux_umol_N_m2_hr_1 = O2_flux_value,
  Ar_flux_umol_N_m2_hr_1 = Ar_flux_value,
  Flow_L_hr_1 = Flow_Rates_df$Flow_rate_L_hr[34:66],
  Area_m2 = Area_m2_value
)
```

## What You Probably Want to See
```{r}
#MIMS Calculations
View(MIMS_fresh_Calc)
View(MIMS_salt_Calc)

#Flux Calculations
View(Flux_Fresh)
View(Flux_Salt)

#Honorable Mentions
View(umoles_nmoles_liter)
View(Flow_Rate_df)
```

# Next steps: verification

```{r}
# basically, we'll want to compare the outputs of our Excel vs. R workflow. I'm picturing a fancy coding method to compare the values and say they are equal (with YHR and MRW datasets)

#Creating a function
validate_flux <- function(flux_df, 
                          excel_file,
                          sheet_name,
                          data_range,
                          tolerance = 1e-6) {
#Tolerance because of floating-point rounding differences 
  
#Loading in Excel Flux Calculations
Excel_Output <- read_excel(excel_file,
  sheet = sheet_name,
  range = data_range)

#Cleaning up Excel
Excel_comp <- Excel_Output %>%
  fill(`CV(N2/Ar)`, `CV(O2/Ar)`, `CV (Ar)`) %>%
  slice(-c(1,2,3,4),) %>% # Removing BP rows (ADJUST)
  filter(!is.na(`N2 flux (umol N m-2 hr-1)`)) %>%
  rename(
    CV_N2_Ar = `CV(N2/Ar)`,
    CV_O2_Ar = `CV(O2/Ar)`,
    CV_Ar = `CV (Ar)`,
    N2_flux_umol_N_m2_hr_1 = `N2 flux (umol N m-2 hr-1)`,
    O2_flux_umol_N_m2_hr_1 = `O2 flux (umol O2 m-2 hr-1)`,
    Ar_flux_umol_N_m2_hr_1 = `Ar flux (umol Ar m-2 hr-1)`,
    Flow_L_hr_1 = `Flow_L_hr-1`,
    Area_m2 = `Area_m2`  ) # <-- Renaming Rows 

#Cleaning up R
Flux_comp <- flux_df%>%
  mutate(
    CV_N2_Ar = as.numeric(gsub("%", "", CV_N2_Ar)),
    CV_O2_Ar = as.numeric(gsub("%", "", CV_O2_Ar)),
    CV_Ar = as.numeric (gsub("%", "", CV_Ar)))
Flux_comp <- Flux_comp %>%
  mutate(across(c(CV_N2_Ar, CV_O2_Ar, CV_Ar), ~ . /100)) # <-- Undoing our previous percentage multiplication (x100) to match Excel

#Columns
cols <- c(
  "CV_N2_Ar",
  "CV_O2_Ar",
  "CV_Ar",
  "N2_flux_umol_N_m2_hr_1",
  "O2_flux_umol_N_m2_hr_1",
  "Ar_flux_umol_N_m2_hr_1",
  "Flow_L_hr_1",
  "Area_m2"
)


# Note: Make sure Column structure/order Matches between R and Excel
Excel_mat <- Excel_comp %>%
  select(all_of(cols)) %>%
  mutate(across(everything(), as.numeric)) %>%
  as.matrix() 
  
R_mat <- Flux_comp %>%
  select(all_of(cols)) %>%
  mutate(across(everything(), as.numeric)) %>%
  as.matrix()

#Comparison
comparison <- round(R_mat,2) - round(Excel_mat,2)

pass_matrix <- abs(comparison) < tolerance

#Putting it into a Data Frame
validation <- data.frame(
  Sample = Flux_comp$sample,
  CV_N2_Ar_Pass = pass_matrix [, 1],
  CV_O2_Ar_Pass = pass_matrix [, 2],
  CV_Ar_Pass = pass_matrix [, 3],
  N2_Pass = pass_matrix [, 4],
  O2_Pass = pass_matrix [, 5],
  Ar_Pass = pass_matrix [, 6],
  Flow_Pass = pass_matrix [, 7],
  Area_Pass = pass_matrix [, 8]
)

validation$Overall_Pass <- apply(validation[-1], 1, all)


#Summary
all_match <- all(validation$Overall_Pass)
if (all_match) {
  cat("Validation PASSED: All samples matched Excel within tolerance.\n")
} else {
  cat("Validation FAILED: Some samples differ from Excel.\n")
}

return(validation)
}


#Inputs
validate_flux(
  flux_df = Flux_Fresh,
  excel_file = "~/Downloads/ECU/Research/Peralta_Lab/Coding_Project/data/ReOrg_Fluxes_4 - Copy.xlsx",
  sheet_name = "May5_Fresh",
  data_range = "X1:AE48" #CV, Flux, Flow, and Area data
)


```

